---
title: 'Homework Assignment #1'
author: "Hornella Fokem Fosso & Jan Lennarz"
date: "1/23/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```



## Exercise 1

In this exercise, it will be simulated for different values of $n$, a path of length $T$ from the VAR($1$) model $$\mathbf{X}_t = A_n\mathbf{X}_{t-1} + \epsilon_t, ~~~ t \in \mathbb{Z},$$ where $(\epsilon_t)  \subset \mathbb{R}$ is s a multivariate standard Gaussian white noise.

```{r include=FALSE}
require(MTS)
set.seed(2301)
```

we perform the simulation for $T=100$ and plot the values of the squared norm corresponding to each $n$.

```{r echo=TRUE, message=FALSE, warning=FALSE}
ns <- c(2,5,10,20) # different values of n
lT <- 100          # length T
sNorm <- numeric(length(ns))  # for storing the squared norms
for (s in 1:length(ns)) {
  n <- ns[s]
  X=matrix(0,nrow=n ,ncol=lT)
  An <- 0.5*diag(n)
  for (i in 1:(n-1)) {
    An[i,i+1] <- 1/5
  }
  for (j in 2:lT) {
    X[,j] <- An %*% X[,(j-1)] + t(t(rnorm(n)))
  }
  Estim <- VAR(t(X),1, output = F)
  A_hat <- Estim$Phi
  B <- A_hat - An
  B <- crossprod(B)
  sNorm[s] <- max(abs(eigen(B)$values))
}

plot(ns, sNorm, xlab = "n", ylab = "squared norm", main = "graph for T = 100")

```

The plot shows that the squared norm increases when n becomes larger. In other words, the estimation error of the  seems to be larger when the dimension of the vector is increased. Let's repeat the experiment for $T=500$ and for $T=1000$

```{r }
ns <- c(2,5,10,20)
lT <- 500
sNorm <- numeric(length(ns))
for (s in 1:length(ns)) {
  n <- ns[s]
  X=matrix(0,nrow=n ,ncol=lT)
  An <- 0.5*diag(n)
  for (i in 1:(n-1)) {
    An[i,i+1] <- 1/5
  }
  for (j in 2:lT) {
    X[,j] <- An %*% X[,(j-1)] + t(t(rnorm(n)))
  }
  Estim <- VAR(t(X),1, output = F)
  A_hat <- Estim$Phi
  B <- A_hat - An
  B <- crossprod(B)
  sNorm[s] <- max(abs(eigen(B)$values))
}

plot(ns, sNorm, xlab = "n", ylab = "squared norm", main = "graph for T = 500")

```



```{r }
ns <- c(2,5,10,20)
lT <- 1000
sNorm <- numeric(length(ns))
for (s in 1:length(ns)) {
  n <- ns[s]
  X=matrix(0,nrow=n ,ncol=lT)
  An <- 0.5*diag(n)
  for (i in 1:(n-1)) {
    An[i,i+1] <- 1/5
  }
  for (j in 2:lT) {
    X[,j] <- An %*% X[,(j-1)] + t(t(rnorm(n)))
  }
  Estim <- VAR(t(X),1, output = F)
  A_hat <- Estim$Phi
  B <- A_hat - An
  B <- crossprod(B)
  sNorm[s] <- max(abs(eigen(B)$values))
}

plot(ns, sNorm, xlab = "n", ylab = "squared norm", main = "graph for T = 1000")

```




```{r}

```
